---
title: "EOF data processing of survey data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(data.table)
```

## Spatial Footprints

```{r EPUareasSurvey, echo = FALSE, eval = TRUE}
EPU <- c('MAB', 'GB', 'GOM')
MAB <- c(1010:1080, 1100:1120, 1600:1750, 3010:3450, 3470, 3500, 3510)
GB <- c(1090, 1130:1210, 1230, 1250, 3460, 3480, 3490, 3520:3550)
GOM <- c(1220, 1240, 1260:1290, 1360:1400, 3560:3830)
NEUSStrata <- c(MAB, GB, GOM)

# plot the EPU areas from the shapefile
library(sf)
ggplot2::ggplot() +
  ggplot2::geom_sf(
    data = NEFSCspatial::BTS_Strata,
    color = "grey",
    alpha = 0.5
  ) +
  ggplot2::geom_sf(
    data = NEFSCspatial::BTS_Strata %>% dplyr::filter(STRATA %in% GOM),
    color = "green"
  ) +
  ggplot2::geom_sf(
    data = NEFSCspatial::BTS_Strata %>% dplyr::filter(STRATA %in% GB),
    color = "red"
  ) +
  ggplot2::geom_sf(
    data = NEFSCspatial::BTS_Strata %>% dplyr::filter(STRATA %in% MAB),
    color = "blue"
  ) +
  ggplot2::coord_sf(xlim = c(-80, -65), ylim = c(35, 46)) +
  ggplot2::theme_minimal() +
  ggplot2::labs(
    title = paste0("EPUs, (", paste0(EPU, collapse = ", "), ") defined by survey strata")
  )


```


## Survey Data

The survey data are from the Northeast Fisheries Science Center (NEFSC) Bottom Trawl Survey (BTS). The BTS is conducted annually in the spring and fall in the Gulf of Maine, Georges Bank, and Mid-Atlantic Bight. The survey data are used to estimate the biomass of fish and invertebrates in the Northeast U.S. continental shelf ecosystem. Data are pulled and swept area estimates are calculated using the [`survdat`](https://noaa-edab.github.io/survdat/) package

```{r surveyData, echo = FALSE, eval = TRUE}
survey <- readRDS(here::here("EDAB/eof/swept_area_biomass_species_EOF.rds"))

survey |> 
  dplyr::group_by(YEAR, EPU) |>
  dplyr::summarise(mt = sum(mt, na.rm = TRUE), .groups = "drop") |>
  ggplot2::ggplot() +
  ggplot2::geom_line(ggplot2::aes(x = YEAR, y = mt, color = EPU)) +
  ggplot2::labs(y = "Swept Area Biomass mt") +
  ggplot2::ggtitle("Total swept area biomass by EPU")
# survey |>
#   ggplot2::ggplot() +
#   ggplot2::geom_line(ggplot2::aes(x = YEAR, y = mt, color = EPU)) +
#   ggplot2::labs(y = "Swept Area Biomass mt") +
#   ggplot2::ggtitle("Total swept area biomass by EPU")

# sum ovr EPUS to obtain total swept area biomass
survey |>
  dplyr::group_by(YEAR) |>
  dplyr::summarise(mt = sum(mt, na.rm = TRUE), .groups = "drop") |>
  ggplot2::ggplot() +
  ggplot2::geom_line(ggplot2::aes(x = YEAR, y = mt)) +
  ggplot2::labs(y = "Swept Area Biomass mt") +
  ggplot2::ggtitle("Total swept area biomass across all EPUs")

```

The bottom trawl survey under samples non demersal species. This will result in an under estimate of system wide biomass. Look at swept area estimates for target species only, .ie. demersal species. Below is a list of the species used to define the demersal species. The list is based on the RPATH groups defined in `ecodata::species_groupings`.

```{r surveyDataDemerals, echo = FALSE, eval = TRUE}
demersal_species <- c("Demersals",
                      "SpinyDogfish",
                      "OtherSkates",
                      "Barndoor",
                      "WinterSkate",
                      "LittleSkate",
                      "Cod",
                      "Haddock",
                      "Pollock",
                      "AmPlaice",
                      "SummerFlounder",
                      "Fourspot",
                      "YTFlounder",
                      "WinterFlounder",
                      "WitchFlounder",
                      "Windowpane",
                      "SmFlatfishes",
                      "BlackSeaBass",
                      "Goosefish",
                      "Redfish",
                      "SmoothDogfish",
                      "OtherFlatfish")

exclude_species <- survey |> 
  dplyr::distinct(RPATH) |>
  dplyr::pull() |> 
  dplyr::setdiff(demersal_species)

# create a gt table of the demrsal species and the species that are excluded
max_length <- max(length(demersal_species), length(exclude_species))
vec1 <- c(demersal_species, rep(NA, max_length - length(demersal_species)))
vec2 <- c(exclude_species, rep(NA, max_length - length(exclude_species)))

species <- data.frame(Demersals = vec1, other = vec2)

gt::gt(species)

```

```{r surveyDataDemeralsPlot, echo = FALSE, eval = TRUE}
survey |>
  dplyr::filter(RPATH %in% demersal_species) |>
  dplyr::group_by(YEAR, EPU) |>
  dplyr::summarise(mt = sum(mt, na.rm = TRUE), .groups = "drop") |>
  ggplot2::ggplot() +
  ggplot2::geom_line(ggplot2::aes(x = YEAR, y = mt, color = EPU)) +
  ggplot2::labs(y = "Swept Area Biomass mt") +
  ggplot2::ggtitle("Total swept area biomass of demersal species by EPU")

survey |>
  dplyr::filter(RPATH %in% demersal_species) |>
  dplyr::group_by(YEAR) |>
  dplyr::summarise(mt = sum(mt, na.rm = TRUE), .groups = "drop") |>
  ggplot2::ggplot() +
  ggplot2::geom_line(ggplot2::aes(x = YEAR, y = mt)) +
  ggplot2::labs(y = "Swept Area Biomass mt") +
  ggplot2::ggtitle("Total swept area biomass of demersal species")



```


