---
title: "Caveats"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(data.table)
```

## Spatial Footprints

```{r EPUsSA, echo = F, eval = TRUE}
gom <- data.table(AREA = c(500, 510, 512:515), EPU = 'GOM')
gb <- data.table(AREA = c(521:526, 551, 552, 561, 562), EPU = 'GB')
mab <- data.table(
  AREA = c(537, 539, 600, 612:616, 621, 622, 625, 626, 631, 632),
  EPU = 'MAB'
)
ss <- data.table(AREA = c(463:467, 511), EPU = 'SS')

epuAreas <- rbindlist(list(gom, gb, mab, ss))
epuAreas[, NESPP3 := 1]
epuAreas[, MeanProp := 1]

# plot the EPU areas from the shapefile
NEFSCspatial::Statistical_Areas_2010_withNames |>
  dplyr::left_join(epuAreas, by = c("Id" = "AREA")) |>
  #dplyr::filter(Id %in% c(gom$AREA, gb$AREA, mab$AREA, ss$AREA)) |>
  ggplot2::ggplot() +
  ggplot2::geom_sf(ggplot2::aes(fill = EPU)) +
  ggplot2::coord_sf(xlim = c(-80, -65), ylim = c(35, 46)) +
  ggplot2::theme_minimal() +
  ggplot2::labs(title = "EPUs defined by Statistical Areas")
```
```{r EPUareasSurvey, echo = FALSE, eval = TRUE}
EPU <- c('MAB', 'GB', 'GOM')
MAB <- c(1010:1080, 1100:1120, 1600:1750, 3010:3450, 3470, 3500, 3510)
GB <- c(1090, 1130:1210, 1230, 1250, 3460, 3480, 3490, 3520:3550)
GOM <- c(1220, 1240, 1260:1290, 1360:1400, 3560:3830)
NEUSStrata <- c(MAB, GB, GOM)

# plot the EPU areas from the shapefile
library(sf)
ggplot2::ggplot() +
  ggplot2::geom_sf(
    data = NEFSCspatial::BTS_Strata,
    color = "grey",
    alpha = 0.5
  ) +
  ggplot2::geom_sf(
    data = NEFSCspatial::BTS_Strata %>% dplyr::filter(STRATA %in% GOM),
    color = "green"
  ) +
  ggplot2::geom_sf(
    data = NEFSCspatial::BTS_Strata %>% dplyr::filter(STRATA %in% GB),
    color = "red"
  ) +
  ggplot2::geom_sf(
    data = NEFSCspatial::BTS_Strata %>% dplyr::filter(STRATA %in% MAB),
    color = "blue"
  ) +
  ggplot2::coord_sf(xlim = c(-80, -65), ylim = c(35, 46)) +
  ggplot2::theme_minimal() +
  ggplot2::labs(
    title = paste0("EPUs, (", paste0(EPU, collapse = ", "), ") defined by survey strata")
  )


```


## Landings and Discards

Landings data are from the commercial fishery and include all species found in CAMS. Discards are estimated from observer data and are only species found in the landings data. Discards are classified as thrown overboard. It does NOT assume all fish are dead. Would need to apply mortality coefficients, especially for elasmobranchs to account for surviving discards. In leiu of this, discards are assumed to be all fish other than elasmobranchs and lobster. Data are pulled and processed using the [`comlandr`](https://noaa-edab.github.io/comlandr/) package

```{r allLandings, echo = FALSE, eval = TRUE}
land <- readRDS(here::here("EDAB/eof/comlandEOFLiveAggGear.rds"))$comland
land |>
  dplyr::filter(EPU %in% c("GOM", "GB", "MAB")) |>
  dplyr::group_by(YEAR, EPU) |>
  dplyr::summarise(mt = sum(SPPLIVMT, na.rm = TRUE), .groups = "drop") |>
  ggplot2::ggplot() +
  ggplot2::geom_line(ggplot2::aes(x = YEAR, y = mt, color = EPU)) +
  ggplot2::labs(y = "Landings mt") +
  ggplot2::ggtitle("Total landings by EPU")

```

```{r otherfishDiscards, echo = FALSE, eval = TRUE}
ofd <- readRDS(here::here("EDAB/eof/otherfish_discards_EOF.rds"))
ofd |>
  dplyr::filter(EPU %in% c("GOM", "GB", "MAB")) |>
  dplyr::group_by(YEAR, EPU) |>
  dplyr::summarise(mt = sum(DISMT, na.rm = TRUE), .groups = "drop") |>
  ggplot2::ggplot() +
  ggplot2::geom_line(ggplot2::aes(x = YEAR, y = mt, color = EPU)) +
  ggplot2::labs(y = "Discards mt") +
  ggplot2::ggtitle("Other fish discards by EPU")

```

Total landings and discards (all fish except elasmobranchs and lobster) across all 3 EPUs

```{r totalLandings, echo = FALSE, eval = TRUE}

ll <- land |>
  dplyr::group_by(YEAR) |>
  dplyr::summarise(mt = sum(SPPLIVMT), .groups = "drop") |>
  dplyr::mutate(type = "landings")

dd <- ofd |>
  dplyr::group_by(YEAR) |>
  dplyr::summarise(mt = sum(DISMT, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(type = "discards")

rbind(ll, dd) |>
  ggplot2::ggplot() +
  ggplot2::geom_line(ggplot2::aes(x = YEAR, y = mt, color = type)) +
  ggplot2::labs(y = "Landings mt") +
  ggplot2::ggtitle("Total landings and discards")

```
Combined

```{r combinedLandingsDiscards, echo = FALSE, eval = TRUE}
rbind(ll, dd) |>
  dplyr::group_by(YEAR) |>
  dplyr::summarise(mt = sum(mt, na.rm = TRUE), .groups = "drop") |>
  ggplot2::ggplot() +
  ggplot2::geom_line(ggplot2::aes(x = YEAR, y = mt)) +
  ggplot2::labs(y = "Landings mt") +
  ggplot2::ggtitle("Total landings and discards combined")

```

### By EPU

```{r byEPU, echo = FALSE, eval = TRUE}
byEPU <- readRDS(here::here("EDAB/eof/total_landings_discards_by_EPU.rds")) 

byEPU |> 
    ggplot2::ggplot() +
  ggplot2::geom_line(ggplot2::aes(x = YEAR, y = mt, color = type)) +
  ggplot2::facet_wrap(~EPU) +
  ggplot2::labs(y = "Landings mt") +
  ggplot2::ggtitle("Landings and discards by EPU")



```

### Discards

* Discard data are discards of species found in the landings data only.
* Discard data are based on observer data. Discards are classified as thrown overboard. It does NOT assume all fish are dead. Would need to apply some mortality coefficient, especially for elasmobranchs. In lieu of this Discards are separated into 3 groups

  * Elasmobranchs
  * Other fish
  * Lobster

```{r allDiscards, echo = FALSE, eval = TRUE}
discards <- readRDS(here::here("EDAB/eof/true_discards_EOF.rds")) 
discards |>
  dplyr::filter(EPU %in% c("GOM", "GB", "MAB")) |>
  dplyr::group_by(NESPP3, RPATH) |>
  dplyr::summarise(mt = sum(DISMT, na.rm = TRUE), .groups = "drop") |>
  dplyr::arrange(dplyr::desc(mt)) |>
  dplyr::slice_head(n = 20) |>
  ggplot2::ggplot() +
  ggplot2::geom_col(ggplot2::aes(x = factor(RPATH), y = mt)) +
  ggplot2::coord_flip() +
  ggplot2::labs(x = "Top 20", y = "Discard mt")

```

Separate discards by groups

```{r discardsElasmobranchs, echo = FALSE, eval = TRUE}

elasmobranchs <- c(
  "Skates",
  "Rays",
  "Sharks",
  "SpinyDogfish",
  "OtherSkates",
  "LittleSkate",
  "SmoothDogfish",
  "Barndoor",
  "WinterSkate"
)
discards |>
  dplyr::filter(EPU %in% c("GOM", "GB", "MAB"))  |>
  dplyr::filter(RPATH %in% elasmobranchs) |> 
  dplyr::group_by(RPATH) |>
  dplyr::summarise(mt = sum(DISMT, na.rm = TRUE), .groups = "drop") |>
  dplyr::arrange(dplyr::desc(mt)) |>
  dplyr::slice_head(n = 20) |>
  ggplot2::ggplot() +
  ggplot2::geom_col(ggplot2::aes(x = factor(RPATH), y = mt)) +
  ggplot2::coord_flip() +
  ggplot2::labs(x = "Elasmobranchs", y = "Discard mt")

```

```{r discardsOther, echo = FALSE, eval = TRUE}

elasmobranchs <- c(
  "Skates",
  "Rays",
  "Sharks",
  "SpinyDogfish",
  "OtherSkates",
  "LittleSkate",
  "SmoothDogfish",
  "Barndoor",
  "WinterSkate"
)
lobster <- "AmLobster"

all_species <- discards |> dplyr::pull(RPATH) |> unique()

otherfish <- setdiff(all_species, c(elasmobranchs, lobster))

readRDS(here::here("EDAB/eof/true_discards_EOF.rds")) |>
  dplyr::filter(EPU %in% c("GOM", "GB", "MAB"))  |>
  dplyr::filter(RPATH %in% otherfish) |> 
  dplyr::group_by(RPATH) |>
  dplyr::summarise(mt = sum(DISMT, na.rm = TRUE), .groups = "drop") |>
  dplyr::arrange(dplyr::desc(mt)) |>
  dplyr::slice_head(n = 20) |>
  ggplot2::ggplot() +
  ggplot2::geom_col(ggplot2::aes(x = factor(RPATH), y = mt)) +
  ggplot2::coord_flip() +
  ggplot2::labs(x = "Other fish", y = "Discard mt")

```

## Survey Data

The survey data are from the Northeast Fisheries Science Center (NEFSC) Bottom Trawl Survey (BTS). The BTS is conducted annually in the spring and fall in the Gulf of Maine, Georges Bank, and Mid-Atlantic Bight. The survey data are used to estimate the biomass of fish and invertebrates in the Northeast U.S. continental shelf ecosystem. Data are pulled and swept area estimates are calculated using the [`survdat`](https://noaa-edab.github.io/survdat/) package

```{r surveyData, echo = FALSE, eval = TRUE}
survey <- readRDS(here::here("EDAB/eof/swept_area_biomass_EOF.rds"))

survey |>
  ggplot2::ggplot() +
  ggplot2::geom_line(ggplot2::aes(x = YEAR, y = mt, color = EPU)) +
  ggplot2::labs(y = "Swept Area Biomass mt") +
  ggplot2::ggtitle("Total swept area biomass by EPU")

# sum ovr EPUS to obtain total swept area biomass
survey |>
  dplyr::group_by(YEAR) |>
  dplyr::summarise(mt = sum(mt, na.rm = TRUE), .groups = "drop") |>
  ggplot2::ggplot() +
  ggplot2::geom_line(ggplot2::aes(x = YEAR, y = mt)) +
  ggplot2::labs(y = "Swept Area Biomass mt") +
  ggplot2::ggtitle("Total swept area biomass across all EPUs")

```
