---
title: "EWE data comparison"
format: html
editor: visual
---

Comparison of 2014 swept area biomass estimates with estimates in 2024.
For the 2024 estimates both fall and spring swept area biomass estimates are calculated using the `survdat` package.

Code from 2014 could be replicated up to the point where swept area biomass is estimated. From that point on it can not be replicated perfectly. Below are figures showing how closely the 2024 estimates compare with the data previously provided

## 2024 Code

The "new" code has been split into 2 functions. One to pull the data, the other to estimate swep area biomass

<details>
```{r 2024code, echo = F,eval = T}
cat(paste0(readLines(here::here("EwE-menhaden-AndreBuchheister/2024request/Andre_data_pulls_v2_update.R")), collapse="\n"))
```
</details>

## Geographic regions 

```{r map,eval = T, echo = T}
source(here::here("EwE-menhaden-AndreBuchheister/2024request/Andre_data_pulls_v2_update_pulldata.r"))
a <- andreprep()

plot(a$polygon)


```
## Areas of geographic regions

When the area of the geographic regions are calculated from the shape file they do not match the areas listed in the 2014 data set. Not sure why since the underlying trawl strata haven't changed.

* Maybe a different set of strata were used from the code we have been using
* Maybe a different projection was implemented
* Maybe a slightly different set of shapefiles were used

```{r areas, eval = T}
data2014 <- readr::read_csv(here::here("EwE-menhaden-AndreBuchheister/2014request/Buchheister_request_biomass.csv"), show_col_types = F) |> 
  dplyr::distinct(Region,AREA) 
data2024 <- survdat::get_area(a$polygon,"eco")

data2024 |> dplyr::left_join(data2014,by = c("STRATUM"="Region")) |> 
  dplyr::rename(Region = STRATUM,
                d2024 = Area,
                d2014 = AREA) |> 
  DT::datatable()

```


## Comparisons

Depending on the species, the spring biomass is a better estimate of the 2014 request than the fall biomass. For other species the opposite is true.

For example

1. gb.deep region - bluefish is not present in the spring but present in the fall, squid is similar.
2. gb.mid region - other small pelagics are better matched in the spring rather than the fall

While the 2014 data can not be reproduced, the trends look similar. Maybe it is time to send the data to Andre?

::: {.panel-tabset}

### Fall

```{r plotfall, eval = T,echo =TRUE, out.width="100%"}
source(here::here("EwE-menhaden-AndreBuchheister/2024request/compare_requests.r"))
compare_requests("fall")

```



### Spring
```{r plotspring, eval = T,echo =TRUE, out.width="100%"}
compare_requests("spring")

```

:::
